MATH/COMP 562 ASSIGNMENT 2

### 2023-07-04 assignment-2/article.tex
Assignment completed

@@

@@

>>>>

\begin{solution}
	Consider three training points in $\mathbb{R}^2$:
	\[
		S = \{((0,0), -1), ((0,1), +1), ((1,0), +1)\}
	\]
	Let's select a vector $w=(0,0.5)$. Define the decision boundary as the hyperplane where $w \cdot x = 0.5$. This classifier $w$ has a loss of $L_S(w) = \frac{1}{3}$ because it misclassifies the point ((1,0),+1).

	Let $0 < \varepsilon < \frac{1}{2}$, and consider the $\varepsilon$-ball around $w$. For any vector $w'$ within this ball (i.e., $|w - w'| \leq \varepsilon$), we observe that $L_S(w')$ will not be lower than $L_S(w)$. This is because in this ball the 0\textendash{}1 loss is constant.

	On the other hand, choosing $w^\prime=(1,1)$ which defines the decision boundary as the hyperplane where $w^\prime \cdot x = 0.5$. This classifier $w^\prime$ correctly classifies all points in the training set S, and hence, its loss $L_S(w^\prime) = 0$. Therefore, we have $L_S(w^\prime) < L_S(w)$, meaning that the vector $w$ is not a global minimum of $L_S$.

@@

\begin{solution}
	Given the logistic loss function
	\[\ell(w, (x, y)) = \log(1 + \exp(-y\langle w, x \rangle))\]

	The magnitude of the gradient of the loss function with respect to $w$ is
	\begin{align*}
		||\nabla \ell(w, (x, y))|| & = \left|\left| -\frac{y \exp(-y\langle w, x \rangle)}{1+\exp(-y\langle w, x \rangle)}x \right|\right| \\
		                           & \leq || -y (1-\sigma(-y\langle w, x\rangle)) \cdot x ||                                               \\
		                           & \leq ||y|| \cdot |1 - \sigma(-y\langle w,x\rangle)| \cdot ||x||                                       \\
		                           & \leq B
	\end{align*}

	Since $||y||=1$, $||x|| \leq B$ and the sigmoid function $\sigma(x)=\frac{1}{1+\exp(-x)}$ is bounded between $[0,1]$. Thus, the loss function is B-Lipschitz.

	To evaluate smoothness, we look at the second derivative of the loss function with respect to $w$. The Hessian is given by
	\[\nabla^2 \ell(w, (x, y)) = \sigma(-y \langle w,x\rangle) \cdot (1-\sigma(-y \langle w,x\rangle))  x x^T \]

	As before, since the sigmoid is bounded in $[0,1]$ the magnitude of $\nabla^2 \ell(w, (x, y))$ is bounded above by $||xx^T|| \leq B^2$.

	Therefore, the logistic regression loss function is both B-Lipschitz and B²-smooth. Note also that the hessian is positive semi-definite, therefore the logistic regression loss is also convex.
\end{solution}

@@

	To show that the hinge loss function is Lipschitz continuous, suppose $w, w' \in \mathbb{R}^d$ and $(x, y) \in X \times Y$, then: \begin{align*}
		| \ell(w, (x, y)) - \ell(w', (x, y)) | & = | \max\{0, 1 - y\langle w, x \rangle\} - \max\{0, 1 - y\langle w', x \rangle\} |                       \\
		                                       & \leq | y\langle w, x \rangle - y\langle w', x \rangle | \quad \text{(since max is Lipschitz with $L=1$)} \\
		                                       & = | y\langle w - w', x \rangle |                                                                         \\
		                                       & \leq \| y(x) \|_2 \| w - w' \|_2 \quad \text{(by Cauchy-Schwarz inequality)}                             \\
		                                       & \leq R \| w - w' \|_2 \quad \text{(since $\| x \|_2 = R$)}
	\end{align*}

	Therefore, the hinge loss function $\ell$ is $R$-Lipschitz.
\end{solution}
